paper3

### 2022-09-19 paper3.Rmd
9/16

# Save the split information for an 80/20 split of the data
split <- landmap %>% 
  filter(!is.na(ScenicQuality)) %>% 
  initial_split(prop = 0.80, strata = ScenicQuality)

@@

```{python plot}
X =  r.landmap.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
# -- play shap
explainer = shap.TreeExplainer(model)

@@

## map together
setwd("Z:\\R\\R projects\\LCA-GB\\figures")
png(filename = "SHAP_values.png", w = 10*2, h = 6*2, units = "in", res = 300)
par(mar = c(0,0,0,0))
multiplot(list(f[1], f[2], f[3], f[4], 
               f[5], f[6], f[7], f[8], 
               f[9], f[10], f[11], f[12]), cols = 4)
dev.off()
```
# Change of support (to 1 km grid)
```{r}
# -- download the administrative region for Great Britain from GADM
UK <- raster::getData('GADM', country = 'GBR', level = 0) |>
  as('sf') |>
  # filter(!NAME_1=='Northern Ireland') |>
  st_transform(crs = 27700)
plot(UK)

# -- download OSGB 1 km grid from github and covert its coordinate system to 27700
grid_1km <- "https://raw.githubusercontent.com/charlesroper/OSGB_Grids/master/GeoJSON/OSGB_Grid_1km.geojson" |>
  geojsonio::geojson_read(what = "sp") |>
  as("sf") |>
  st_transform(crs = 27700)

@@

# -- spatial join OSGB 1 km grid with Scenic-Or-Not data
test <- st_join(grid_1km, sc)

test1 <- test |>
  aggregate(cbind(ID,
                  `Geograph URI`,
                  Votes,
                  Number,
                  Mean,
                  Median,
                  BayeScore,
                  Entropy,
                  Variance)~PLAN_NO, data = ., paste0, collapse = ",") %>%
  merge(LANDMAP, ., by = "UID", all = T) %>%
  as_tibble()

plot(UK['NAME_0']) +
  plot(test[is.na(test$ID),'Mean'])

grid_1km <- st_join(grid_1km, sc, join = st_within, left = TRUE) |>
  dplyr::select(PLAN_NO, 
                `Scenic-Or-Not ID` = ID,
                `Geograph URI`,
                Votes,
                Number,
                Mean,
                Median,
                BayeScore,
                Entropy,
                Variance)

grid_1km <- st_transform(grid_1km, crs = st_crs(Wildness)) %>%
  exact_extract(Wildness, ., 
                fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), 
                quantiles = c(0.25, 0.75)) |>
  mutate(Wildness.mean = mean,
         Wildness.median = median,
         Wildness.iqr = q75 - q25,
         Wildness.cv = coefficient_of_variation,
         Wildness.variance = variance) |>
  dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) |>
  bind_cols(grid_1km, .) |>
  st_transform(crs = 27700)

```

>>>>

# Save the split information for an 80/20 split of the data
split <- landmap_B %>% 
  filter(!is.na(ScenicQuality)) %>% 
  initial_split(prop = 0.80, strata = ScenicQuality)

@@

```{python plot}
X =  r.landmap.loc[:, ['Scenicness.weighted_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
# -- play shap
explainer = shap.TreeExplainer(model)

@@

## map together
setwd("Z:\\R\\R projects\\LCA-GB\\figures")
png(filename = "SHAP_values.png", w = 10*2, h = 6*2, units = "in", res = 300)
par(mar = c(0,0,0,0))
multiplot(list(f[1], f[2], f[3], f[4], 
               f[5], f[6], f[7], f[8], 
               f[9], f[10], f[11], f[12]), cols = 4)
dev.off()
```
# Change of support (to 1 km grid)
```{r}
# -- download OSGB 1 km grid from github and covert its coordinate system to 27700
grid_1km <- "https://raw.githubusercontent.com/charlesroper/OSGB_Grids/master/GeoJSON/OSGB_Grid_1km.geojson" |>
  geojsonio::geojson_read(what = "sp") |>
  as("sf") |>
  st_transform(crs = 27700)

@@

# -- spatial join OSGB 1 km grid with Scenic-Or-Not data for the sake of understanding the distribution of Scenic-Or-Not ratings

# grid_1km <- st_join(grid_1km, sc) |>
#   st_drop_geometry() |>
#   group_by(PLAN_NO) |>
#   summarise(ID = paste(ID, collapse = ','),
#             `Geograph URI` = paste(`Geograph URI`, collapse = ','),
#             Votes = paste(Votes, collapse = ','),
#             Number = paste(Number, collapse = ','),
#             Mean = paste(Mean, collapse = ','),
#             Median = paste(Median, collapse = ','),
#             BayeScore = paste(BayeScore, collapse = ','),
#             Entopy = paste(Entropy, collapse = ',')) |>
#   rowwise() |>
#   mutate(Photos = ifelse(ID == 'NA', 0, strsplit(ID, ',') %>% unlist() %>% length())) |>
#   ungroup() %>%
#   left_join(grid_1km, ., by = c('PLAN_NO' = 'PLAN_NO'))

grid_1km <- st_join(grid_1km, sc) %>%
  aggregate(cbind(BayeScore, Number, VR = BayeScore*Number)~PLAN_NO, data = ., paste0, collapse = ',') %>%
  merge(grid_1km, ., by = 'PLAN_NO', all = T) |>
  st_drop_geometry() |>
  rowwise() |>    
  mutate(Scenicness.number_of_all = strsplit(Number, ',') %>% unlist() %>% as.numeric() %>% sum(na.rm = TRUE),
         Scenicness.mean_of_bayescore = strsplit(BayeScore, ',') %>% unlist() %>% as.numeric() %>% mean(),
         Scenicness.median_of_bayescore = strsplit(BayeScore, ',') %>% unlist() %>% as.numeric() %>% median(),
         Score_multiplied_by_Number = strsplit(VR, ',') %>% unlist() %>% as.numeric() %>% sum(na.rm = TRUE),
         Scenicness.weighted_bayescore =  Score_multiplied_by_Number/(Scenicness.number_of_all + 3) + 
                                          3*Scenicness.mean_of_bayescore/(Scenicness.number_of_all + 3)) |>
  dplyr::select(-c(BayeScore, Number, VR, Score_multiplied_by_Number, Scenicness.number_of_all)) |>
  ungroup() %>%
  left_join(grid_1km, ., by = c('PLAN_NO' = 'PLAN_NO'))

# -- extract raster values of wildness quality based on 1 km OSGB grid 
library(exactextractr)

st_crs(Wildness) == st_crs(grid_1km)

grid_1km <- st_transform(grid_1km, crs = st_crs(Wildness)) %>%
  exact_extract(Wildness, ., 
                fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), 
                quantiles = c(0.25, 0.75)) |>
  mutate(Wildness.mean = mean,
         Wildness.median = median,
         Wildness.iqr = q75 - q25,
         Wildness.cv = coefficient_of_variation,
         Wildness.variance = variance) |>
  dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) %>%
  bind_cols(grid_1km, .) |>
  st_transform(crs = 27700)

# -- calculate the centroid coordinate for each grid cell
grid_1km <- mutate(grid_1km, 
                   x_coord = st_centroid(grid_1km$geometry) %>% st_coordinates() %>% .[, 1],
                   y_coord = st_centroid(grid_1km$geometry) %>% st_coordinates() %>% .[, 2])

# -- download the administrative region for Great Britain from GADM for subseting the 1 km grid
GB <- raster::getData('GADM', country = 'GBR', level = 1) |>
  as('sf') |>
  filter(!NAME_1=='Northern Ireland') |>
  st_union() |>
  st_transform(crs = 27700)
plot(GB)

grid_1km <- grid_1km[GB, ]
```
# prepare environment for running python
```{r}
Sys.setenv("RETICULATE_PYTHON" = ".\\.venv\\Scripts\\python.exe")
# Sys.getenv()
Sys.which("Python")

# SHAP value
```{python plot}
import shap

X =  r.grid_1km.loc[:, ['Scenicness.weighted_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
# -- play shap
explainer = shap.TreeExplainer(model)

# landmap = r.landmap

# X_vars = ['Scenicness','Wildness', 'x_coord', 'y_coord']
# 
# import matplotlib.pyplot as plt
# 
# fig, ax = plt.subplots(2,2,figsize=(15,15))
# 
# for j in range(len(shap_values)):
#     ax = ax.ravel()
#     # X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
#     #              vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)
#     fig[j] = shap.summary_plot(shap_values[1],
#                               X.values,
#                               feature_names = X_vars,
#                               show = False),
#                               # plot_type = 'compact_dot', # making the dots disappear
#                               plot_size = (50, 6))
#              plt.gcf().axes[-1].set_aspect('auto')
#              plt.tight_layout()
#              plt.gcf().axes[-1].set_box_aspect(40)
#     ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)
#     
# fig.delaxes(ax[-1])
# 
# 
# shap.summary_plot(shap_values[1],
#                   X.values,
#                   feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], show = False),
#                   # plot_type = 'compact_dot', # making the dots disappear
#                   plot_size = (50, 6))
# plt.gcf().axes[-1].set_aspect('auto')
# plt.tight_layout()
# plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
# plt.show(); plt.close()
# plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()
```
## Spatial patterns of SHAP values
```{r}
shap_values = py$shap_values
shap_Low = data.frame(shap_values[[1]]) %>% transmute(SHAP_Scenicness_Low = X1, 
                                                      SHAP_Wildness_Low = X2,
                                                      SHAP_Location_Low = X3 + X4) 
shap_Moderate = data.frame(shap_values[[2]]) %>% transmute(SHAP_Scenicness_Moderate = X1,
                                                           SHAP_Wildness_Moderate = X2,
                                                           SHAP_Location_Moderate= X3 + X4)
shap_High = data.frame(shap_values[[3]]) %>% transmute(SHAP_Scenicness_High = X1,
                                                       SHAP_Wildness_High = X2,
                                                       SHAP_Location_High = X3 + X4)
shap_Outstanding = data.frame(shap_values[[4]]) %>% transmute(SHAP_Scenicness_Outstanding = X1,
                                                              SHAP_Wildness_Outstanding = X2,
                                                              SHAP_Location_Outstanding = X3 + X4)

grid_1km_shap = cbind(grid_1km, shap_Low, shap_Moderate, shap_High, shap_Outstanding)
rm(list = c('shap_Low', 'shap_Moderate', 'shap_High', 'shap_Outstanding'))
```
## plotting the spatial distribution of SHAP values
```{r}
library(ggplot2)

f2 <- list()

for (i in 1:12) {
  f2[[i]] <- ggplot() +
    geom_sf(data = grid_1km_shap, aes(fill = !!sym(names(grid_1km_shap)[i+11])), colour = NA) +
    theme_map() +
    labs(x = NULL, 
         y = NULL, 
         title = names(grid_1km_shap)[i+11]) + # stringr::str_c("SHAP for ", collapse = " ")
    theme(legend.position = "bottom",
          title = element_text(size = 15)) +
    scale_fill_gradient2(low = ("#CB181D"),
                         mid = "white",
                         name = NULL,
                         high = ("#2171B5"),
                         midpoint = 0,
                         space = "Lab",
                         na.value = "white",
                         guide = guide_colourbar(title.position = "top",
                                                 # title.hjust = 0.5,
                                                 title.vjust = 0.8,
                                                 barwidth = 15))
}

setwd("Z:\\R\\R projects\\LCA-GB\\figures")
png(filename = "SHAP_values_4.png", w = 10*2, h = 6*2, units = "in", res = 300)
par(mar = c(0,0,0,0))
multiplot(list(f2[1], f2[2], f2[3], f2[4], 
               f2[5], f2[6], f2[7], f2[8], 
               f2[9], f2[10], f2[11], f2[12]), cols = 4)
dev.off()
```
